---
title: "Counterstrike"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: aylmer.bib
---



```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("aylmer")
```

To cite the `aylmer` package in publications, please use @west2008.
The data is taken from an influential match between "FaZe Clan" and
"Cloud9", two of the most successful E-sports syndicates of all time,
when they competed at Boston 2018.  More background information on the
dataset and its source is given at `counterstrike.Rd` of the `hyper2`
package [@hankin2017_rmd].  This short document discusses a dataset
taken from Boston 2018.  It was provided by Zachary Hankin.


```{r zacskills}
counterstrike_kills <- matrix(c(
    0, 4, 4, 6, 0,10, 0, 0, 0, 3, 
    7, 0, 0, 0, 6, 0, 4, 4, 7, 0, 
    9, 0, 0, 0, 5, 0, 7, 5, 9, 0, 
    5, 0, 0, 0, 6, 0,10, 8, 2, 0, 
    0, 7, 3, 5, 0, 6, 0, 0, 0, 6, 
    7, 0, 0, 0, 7, 0, 8, 4, 6, 0, 
    0, 4, 2, 4, 0, 6, 0, 0, 0, 8, 
    0, 6,12, 4, 0, 3, 0, 0, 0, 3, 
    0, 4, 3, 3, 0, 3, 0, 0, 0, 4, 
    0, 0, 0, 0, 4, 0, 2, 6, 6, 0
), 10, 10, byrow=TRUE)

players <- 
    c("NiKo","autimatic","tarik","Skadoodle","olofmeister", 
      "Stewie2k", "karrigan", "GuardiaN","rain","RUSH" )
      

## rearrange rows and columns so that rows 1-5 are team 1 and rows
## 6-10 are team 2:
## team1: autimatic,tarik, Skadoodle,Stewie2k,RUSH
## team2: NiKo,olofmeister,karrigan,GuardiaN,rain

dimnames(counterstrike_kills) <-
    list(
        killer = players,
        victim = players
    )

jj <- c(
    2,3,4,6,10,   # team 1
    1,5,7,8,9     # team 2
)

counterstrike_kills <- counterstrike_kills[jj,jj]

## Because shooting members of your own team is forbidden, create
## structural zeros

jj <- as.matrix(expand.grid(1:5,1:5))
counterstrike_kills[jj  ,jj  ] <- NA
counterstrike_kills[jj+5,jj+5] <- NA


counterstrike_kills
```


OK now the original thinking was that the structural zeros indicated
the Aylmer test.  But on reflection, any member of one team can kill
any member of the other.  So it might be better to consider the "kill
matrix" with rows being team A, columns being team B, and entry
$(i,j)$ being number of times player $i$ of team A kills player $j$ of
team B.  We have symmetry between the teams so can create two
matrices:


```{r}
(cloud9_kills_faze <- counterstrike_kills[1:5 ,6:10])
(faze_kills_cloud9 <- counterstrike_kills[6:10,1:5 ])
```


Now test for homogeneity of proportions between the rows using Fisher:

```{r}
fisher.test(cloud9_kills_faze, simulate.p.value = TRUE)
fisher.test(faze_kills_cloud9, simulate.p.value = TRUE)
```


No evidence either way.


# References {-}
