---
title: "Frogs and their mating preferences"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: aylmer.bib
---


```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("aylmer")
```

To cite the `aylmer` package in publications, please use @west2008.
This short document discusses a dataset collected and discussed by
@kirkpatrick2006.  Their table 1, in count form, is:

```{r kirkpatrickfrogcountdata}
"frogs.matrix" <-
  matrix(c(
            NA, 10L,  8L,  7L,  6L,  7L,  4L,  5L,  3L,
           10L,  NA,  7L, 12L,  8L,  4L,  1L,  5L,  4L,
           12L, 13L,  NA,  8L, 10L,  6L,  8L,  8L,  2L,
           13L,  8L, 12L,  NA, 10L, 10L,  4L,  4L,  9L,
           14L, 12L, 10L, 10L,  NA, 15L, 10L,  8L,  2L,
           13L, 16L, 14L, 10L,  5L,  NA,  6L, 11L,  9L,
           16L, 19L, 12L, 16L, 10L, 14L,  NA,  5L,  8L,
           15L, 15L, 12L, 16L, 12L,  9L, 15L,  NA, 13L,
           17L, 16L, 18L, 11L, 18L, 11L, 12L,  7L,  NA
           ), 9, 9, byrow=TRUE)

rownames(frogs.matrix) <- 
c("Sc", "Sb", "Ob", "Oa", "Oc", "Sa", "Sd", "Od", "M")
colnames(frogs.matrix) <- rownames(frogs.matrix)
frogs.matrix
```

Thus `frogs.matrix` is a 9-by-9 matrix with rows and columns
corresponding to 9 stimuli presented to a female frog.  Entry $(i,j)$
corresponds to the number of times the frog preferred stimulus $i$ to
stimulus $j$.  To analyse this with the `aylmer` package we need to
use `as.pairwise()`:

```{r}
(frogs <- as.pairwise(frogs.matrix))
```

Dataset `frogs` is thus a 36-by-9 matrix with each row corresponding
to a pair of distinct stimuli; if a row has non-`NA` columns $i$ and
$j$, then its entries show the number of times the frog preferred
stimulus $i$ or stimulus $j$.


```{r}
aylmer.test(frogs, simulate.p.value = TRUE)
```

### Package dataset {-}

Following lines create `frogs.rda`, residing in the `data/` directory of the package.

```{r,label=savefrogsdataset}
save(frogs, frogs.matrix, file="frogs.rda")
```

### References {-}





