---
title: The Ronald test"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: aylmer.bib
---


```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("aylmer")
```


To cite the `aylmer` package in publications, please use
@west2008.  Here I consider pairwise comparisons with a
home-ground advantage and wish to assess whether all teams have the same degree of home advantage.

# Start simple: two teams {-}


```{r,echo=FALSE}
jj <- matrix(c(NA, 7+2i ,6+5i, NA),2,2,byrow=TRUE)
teams <- LETTERS[1:2]
dimnames(jj) <- list("@home" = teams,"@away"=teams)
dimnames(jj) <- list("@home" = teams,"@away"=teams)
print(jj)
```

Above we see that A played at home 9 times, winning 7 and losing 2
[row 1 column 2 is $7+2i$]; B played at home 11 times, winning 6 and
losing 5 [row 2 column 1 is $6+5i$] (observe that the real components
correspond to home wins and the imaginary components correspond to
away wins.  Thus A wins 7 times at home and 5 times away).  In its
most general form, the $2\times 2$ result matrix would be

$$
\begin{bmatrix}
-&\alpha+\beta i \\
\gamma + \delta i&-\\
\end{bmatrix}
$$


```{r,label=try2x2}
M <- matrix(c(
    NA, 5+1i, 9+2i, NA),
    nrow=2,ncol=2,byrow=TRUE)
teams <- LETTERS[seq_len(2)]
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
M
```

## Four teams

Now a more complicated example, a
modified version of the 4x4 matrix on `home_advantage.Rmd`:


```{r label=definehomeawaymatrix}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

away_games_won <- matrix(c(
    NA, 05, 02, 02,
     9, NA, 10, 02,
     3, 04, NA, 07,
     8, 06, 04, NA),
    nrow=4,ncol=4,byrow=TRUE)

teams <- LETTERS[1:4]
dimnames(home_games_won) <- list("@home" = teams,"@away"=teams)
dimnames(away_games_won) <- list("@home" = teams,"@away"=teams)

D <- home_games_won + 1i*away_games_won
D
```

Above we see, from row 1 column 2 [`36+10i`], that A played at home
against B a total of $36+10=46$ times, winning $36$ and losing $10$;
From row 2 column 1 [`11+10i`] we see that B played at home against A
a total of 21 times, winning 11 and losing 10. 

We will enumerate all arrays with fixed inline sums:

```{r}
no.of.boards(away_games_won)
```

```{r cache=TRUE}
agw <- allboards(away_games_won)  # every possible result with consistent marginals
dim(agw)
hgw <- -sweep(agw, 1:2, home_games_won + away_games_won)
hgw[,,1:3]
wanted <- apply(hgw,3,function(x){all(x >= 0, na.rm=TRUE)})
table(wanted)
```

Above we see that every board congruent to [that is, marginal totals
matching] `away_games_won` (viz `agw`) leads to a non-negative board
for `home_games_won`.

```{r}
test_statistic <- 1/prod(factorial(c(home_games_won,away_games_won)),na.rm=TRUE)
f <- function(x){1/prod(factorial(x),na.rm=TRUE)}
resampled_probs <- apply(agw,3,f) * apply(hgw,3,f)
hist(log(resampled_probs))
plot(log(resampled_probs),cex=0.2)
abline(h = log(test_statistic))
pvalue <- sum(resampled_probs[resampled_probs < test_statistic]) / sum(resampled_probs)
pvalue
```

So for this statistic we have no evidence to reject the null of the
same home advantage for all teams.  Now I will modify the data so
there is a significant difference:

```{r}
home_games_won[] <- 3
away_games_won[] <- 3
diag(home_games_won) <- NA
diag(away_games_won) <- NA
home_games_won[1,3] <- 19
home_games_won[3,1] <- 0
away_games_won[1,3] <- 0
away_games_won[3,1] <- 19
home_games_won
away_games_won
no.of.boards(away_games_won)
```

```{r}
agw <- allboards(away_games_won)  # every possible result with consistent marginals
dim(agw)
hgw <- -sweep(agw, 1:2, home_games_won + away_games_won)
wanted <- apply(hgw,3,function(x){all(x >= 0, na.rm=TRUE)})
table(wanted)
```

Above we see from the table that XXX of the possible tables for
`away_games_won` are forbidden, for they yield negative entries in the
`home_games_won` results.

```{r}
hgw <- hgw[,,wanted]
agw <- agw[,,wanted]
test_statistic <- 1/prod(factorial(c(home_games_won,away_games_won)),na.rm=TRUE)
f <- function(x){1/prod(factorial(x),na.rm=TRUE)}
resampled_probs <- apply(agw,3,f) * apply(hgw,3,f)
hist(log(resampled_probs))
plot(log(resampled_probs),cex=0.2)
abline(h = log(test_statistic))
pvalue <- sum(resampled_probs[resampled_probs < test_statistic]) / sum(resampled_probs)
pvalue
```


## References {-}
